Title

### 2013-04-05 TCAD2013/overclocking_tcad.tex
CSA overall check v2

\subsection{Timing Models for Carry Select Adder}
We initially describe the modelling method for the CSA timing, with the aim of forming the relationship between the operating frequency and the corresponding maximum word-length of CSA. This information can be employed to determine the truncation error based on the models presented in Section.xxx.

In a CSA with $s$ stages~($s\geqslant 2$), let the stage delay be denoted by $d_{0}\dots d_{s-1}$, where $d_{0}$ and $d_{s-1}$ represent the delay of the most significant stage and the least significant stage, respectively. Unlike other stages, the least significant stage is built by one RCA only, since it is directly driven by the carry input. In our analysis, we follow the previous assumption that delay is due to carry propagation as well as, in this case, multiplexing the carry output. Therefore the delay of the $i^{th}$ stage can be obtained through~(\ref{CSA_SingleStageDelay}), where $\mu_{c}$ denotes the delay of $1$-bit carry propagation and $\mu_{mux}$ denotes the delay of multiplexing.
%Note that the delay of the ${(s-1)}^{th}$ and ${(s-2)}^{th}$ stages are identical, because the least significant stage of CSA is built by RCA.
\begin{eqnarray}\label{CSA_SingleStageDelay}
  d_i=\left\{
    \begin{matrix}
      n_i\cdot \mu_{c}+(i+1)\cdot\mu_{mux}, &\textrm{if $i\in\left[0,s-2\right]$}\\
      d_{i-1}, & \textrm{if $i=s-1$}
    \end{matrix}
  \right.
\end{eqnarray}

Under the timing-driven design environment, the delay of each stage of CSA is set to be approximately uniform for the fastest operation. In this case we obtain~(\ref{CSA_DelayEqual}).
\begin{eqnarray}\label{CSA_DelayEqual}
  d_0=d_1=\cdots=d_{s-1}
\end{eqnarray}

Substituting (\ref{CSA_SingleStageDelay}) into (\ref{CSA_DelayEqual}) yields (\ref{CSA_DelayRep}), which denotes the CSA word-length of stage $i$.
%Note that the word-length of stage $s-1$ and $s-2$ are identical, since the least significant stage is composed of RCA.
\begin{eqnarray}\label{CSA_DelayRep}
 n_i=\left\{
	\begin{matrix}
	  n_0-i\cdot\frac{\mu_{mux}}{\mu_c}, & \textrm{if $i\in\left[0,s-2\right]$}\\
	  n_{i-1}, &\textrm{if $i=s-1$}
      %n_0-(c-2)\cdot\frac{\mu_{mux}}{\mu_c}, & \textrm{if $i=c-1$ and $c\geqslant2$}
	\end{matrix}
    \right.
\end{eqnarray}

The word-length of CSA is then given by (\ref{CSA_WL}).
\begin{eqnarray}\label{CSA_WL}
  n_{CSA}=\sum_{i=0}^{s-1}n_{i}=s\cdot n_{0}-\frac{\mu_{mux}}{\mu_{c}}\cdot\frac{(s+1)(s-2)}{2}
\end{eqnarray}

In the conventional situation, the word-length of both RCA and CSA should be properly selected in order to meet timing. In this case,~(\ref{CSA_RCA}) can be obtained, where $n_{RCA}$ is determined by current timing information through (xxx).
\begin{eqnarray}\label{CSA_RCA}
  \mu_{c}\cdot n_{RCA}=\mu_{c}n_0+\mu_{mux}
\end{eqnarray}

Based on~(\ref{CSA_WL}) and~(\ref{CSA_RCA}), we may obtain the representation of the word-length of CSA in terms of a given timing constraint, as presented in~(\ref{CSA_Timing}).
% we form the relationship between the word-length of CSA and RCA under a given timing constraint, as presented in~(\ref{CSA_Timing}).
\begin{eqnarray}\label{CSA_Timing}
  n_{CSA}=s\cdot n_{RCA}-\frac{\mu_{mux}}{\mu_{c}}\cdot\frac{(s+2)(s-1)}{2}
\end{eqnarray}

@@

\subsection{Model Verification}
As seen in (\ref{CSA_Timing}), the value of $\mu_{mux}/\mu_c$ should be determined before applying the timing model. This ratio can be obtained through timing analysis. Specifically, varying the word-length of a single stage $n_i$ and recording the corresponding delay value $d_i$ through the timing analysis tool. According to~(\ref{CSA_SingleStageDelay}), the ratio can be obtained by fitting those values using linear polynomials.
%, as presented in Fig.~(\ref{DelayFitting}). Out experimental results show that $\mu_{mux}/\mu_c\approx8$.
%
%\begin{figure}[htbp]
%  \centering
%  \includegraphics[width=3.3in]{./Figures/Fitting.eps}
%  \caption{Fitted curve of (\ref{CSA_SingleStageDelay}) for the most significant stage ($i=0$), based on the given input word-length and the corresponding delay value obtained through the Xilinx Timing Analyzer.}
%  \label{DelayFitting}
%\end{figure}

Using this information, we verify our timing model with experimental results, which are obtained from post place and route simulations on Xilinx Virtex-6 FPGAs. Fig.~\ref{CSA Model Verification} demonstrates both the modelled value and the experimental results of the maximum word-length of the 2-stage CSA for various operating frequencies. In this experiment, the input data are randomly sampled from a 16-bit data set, which follows uniform distribution. It can be seen that the the modelled value match well with the experimental results.
%The maximum input word-length in our experiments is 16-bit, hence the modelled value is set to 16 if it expires.
\begin{figure}[htbp]
  \centering
  \includegraphics[width=3in]{./Figures/Model.eps}
  \caption{Comparison of the maximum word-length for a given operating frequency between the modelled value and the experimental results from FPGA simulations.}
  \label{CSA Model Verification}
\end{figure}

@@

\subsection{Area Overhead}
Fig.~\ref{CSA3stage Timing} demonstrates the maximum word-lengths for RCA and CSA with 2 stages and 3 stages under a range of operating frequencies. We only investigate 3 stages as the maximum stages number predicted in (\ref{CSA_StageNumber}) is 3. It can be seen that in comparison to RCA, CSA achieves greater word-length when frequency is initially increased. This corresponds to smaller truncation errors. RCA only outperforms than CSA when very high frequency is applied. In addition, the word-length of 3-stage CSA is always greater than 2-stage CSA across the entire frequency domain, as expected.
%\begin{figure}[htbp]
%  \centering
%  \includegraphics[width=3.3in]{./Figures/CSA3stage_Timing.eps}
%  \caption{Maximum word-length for RCA and CSA across a variety of frequencies.}
%  \label{CSA3stage Timing}
%\end{figure}

\begin{figure}[htbp]
    \begin{minipage}[b]{.48\textwidth}
        \centering
        \subfigure[Maximum word-length.]{
            \includegraphics[width=3.3in]{./Figures/CSA3stage_Timing.eps}
            \label{CSA3stage Timing}
        }
        \subfigure[Hardware resource usage.]{
            \includegraphics[width=3.3in]{./Figures/CSA3stage_Area.eps}
            \label{CSA3stage Area}
        }
    \end{minipage}
    \centering
    \caption{Comparison between RCA and CSA across a variety of frequency values in terms of the maximum word-length of input signal and the area consumption.}
\end{figure}

However, the accuracy benefits brought by CSA comes with the cost of large area overhead. Fig.~\ref{CSA3stage Area} depicts the resource usage (in terms of the number of Look-Up Tables (LUTs) in the FPGA) used for all three structures. It can be seen that for a given frequency, the 3-stage CSA consumes $2.4\times\sim3.7\times$ area than RCA, while the number of the 2-stage CSA is $1.7\times\sim3.1\times$. This finding poses a question of whether CSA with higher stage numbers still offer better accuracy than RCA under a limited area budget.
%\begin{figure}[htbp]
%  \centering
%  \includegraphics[width=3.3in]{./Figures/CSA3stage_Area.eps}
%  \caption{Area overhead for RCA and CSA.}
%  \label{CSA3stage Area}
%\end{figure}

\subsection{Exploring Trade-offs Between Accuracy, Performance and Area}
%Although advanced architectures such as CSA inherently offer better performance than the basic structure, this would generate a large area overhead. In the following experiments,
To answer this question, we explore the trade-offs between accuracy, performance as well as area in this section. If the available hardware resources are limited, the full word-length of both CSA and RCA might not be implemented. The precision lose might generates large errors even at low frequencies. This is demonstrated by experiments where area constraints are applied besides the timing requirement. Conventionally, both of the constraints are met by reducing the word-length of the input signal. The errors at the outputs are recorded with the reference to the original word-length, which is 16 bits in the following experiments. In addition, we propose another designing method where RCA is implemented with the maximum possible word-length under the given area budget, while the timing constraints are met by overclocking. For a certain pair of $\{Area, Frequency\}$ constraint, the optimum design method is selected based on the following criteria:
%For instance, Fig.xxx to Fig.xxx illustrate error expectations of all the described design scenarios when setting the LUTs number to 45, 35, 25 and 15 respectively. The optimal design metric with the minimum error expectation is labelled for each area constraint. For a given frequency and area constraint, the optimum design is chosen based on the following metrics:
\begin{itemize}
  \item Design with the minimum error expectation is the optimum design;
  \item If multiple designs achieve the same error expectation, then the design with minimum area is the optimum design;
  \item If the error expectation and area are identical for multiple designs, they are all treated as the optimum design.
\end{itemize}

The optimum design method for a variety of frequency and area constraints is demonstrated in Fig.~\ref{Tradeoff}, from which several observations can be made. If the available area is large enough, CSA serves as the optimum design option because it inherently operates faster than RCA, and it can be implemented with full precision in this situation. The 2-stage CSA is better than the 3-stage CSA when frequency is initially increased, as it consumes less area although both of them achieve the same error expectation. For a tighter area budget, only part of the CSA can be implemented, whilst RCA still keeps full precision. In this case, area instead of timing becomes the dominate factor for small frequency values. This leads to precision loss for CSA initially. Therefore we see RCA with overclocking is the best design across almost the whole frequency domain. For an even stringent area constraint, the word-length of RCA is also limited. This results in truncation error initially for all design scenarios. Meanwhile, CSA with high stage numbers could not even be implemented at all. 
\begin{figure}[htbp]
  \centering
  \includegraphics[width=3.5in]{./Figures/Tradeoff.eps}
  \caption{Demonstration of the optimum design methodology for a variety frequency and area constraints.}
  \label{Tradeoff}
\end{figure}

>>>>

\subsection{Timing Models for Carry Select Adder}
We initially describe the modelling method for the CSA timing, with the aim of forming the relationship between the operating frequency and the corresponding maximum word-length of CSA. This information can then be employed to determine the truncation error based on the models presented in Section.xxx.

In a CSA with $s$ stages~($s\geqslant 2$), let the stage delay be denoted by $d_{0}\dots d_{s-1}$, where $d_{0}$ and $d_{s-1}$ represent the delay of the most significant stage and the least significant stage, respectively. Unlike other stages, the least significant stage is built by one RCA only, since it is directly driven by the carry input. In our analysis, we follow the previous assumption that the critical path delay is due to carry propagation as well as, in this case, multiplexing the carry output. Therefore, the delay of the $i^{th}$ stage can be obtained through~(\ref{CSA_SingleStageDelay}), where $\mu_{c}$ denotes the delay of $1$-bit carry propagation and $\mu_{mux}$ denotes the delay of multiplexing.
%Note that the delay of the ${(s-1)}^{th}$ and ${(s-2)}^{th}$ stages are identical, because the least significant stage of CSA is built by RCA.
\begin{eqnarray}\label{CSA_SingleStageDelay}
  d_i=\left\{
    \begin{matrix}
      n_i\cdot \mu_{c}+(i+1)\cdot\mu_{mux}, &\textrm{if $i\in\left[0,s-2\right]$}\\
      d_{s-2}, & \textrm{if $i=s-1$}
    \end{matrix}
  \right.
\end{eqnarray}

Under the timing-driven design environment, the delay of each stage of CSA is equalized in order to achieve the fastest operation. That is,~(\ref{CSA_DelayEqual}) can be obtained.
\begin{eqnarray}\label{CSA_DelayEqual}
  d_0=d_1=\cdots=d_{s-1}
\end{eqnarray}

This design metric also consumes the maximum word-length of CSA for a given frequency, potentially corresponds to the minimum precision loss at inputs. Substituting (\ref{CSA_SingleStageDelay}) into (\ref{CSA_DelayEqual}) yields (\ref{CSA_DelayRep}), where $n_i$ denotes the CSA word-length of stage $i$.
%Note that the word-length of stage $s-1$ and $s-2$ are identical, since the least significant stage is composed of RCA.
\begin{eqnarray}\label{CSA_DelayRep}
 n_i=\left\{
	\begin{matrix}
	  n_0-i\cdot\frac{\mu_{mux}}{\mu_c}, & \textrm{if $i\in\left[0,s-2\right]$}\\
	  n_{s-2}, &\textrm{if $i=s-1$}
      %n_0-(c-2)\cdot\frac{\mu_{mux}}{\mu_c}, & \textrm{if $i=c-1$ and $c\geqslant2$}
	\end{matrix}
    \right.
\end{eqnarray}

The word-length of CSA is then given by (\ref{CSA_WL}).
\begin{eqnarray}\label{CSA_WL}
  \begin{split}
    n_{CSA} &=\sum_{i=0}^{s-1}n_{i} \\
    &= s\cdot n_{0}-\frac{\mu_{mux}}{\mu_{c}}\cdot\frac{(s+1)(s-2)}{2}
  \end{split}
\end{eqnarray}

Conventionally for a given frequency constraint, the word-length of both RCA and CSA should be properly selected to avoid timing violation, as presented in~(\ref{CSA_RCA}), where $n_{RCA}$ is determined by current timing information through (xxx).
%In the conventional situation, the word-length of both RCA and CSA should be properly selected in order to meet timing. In this case,~(\ref{CSA_RCA}) can be obtained, where $n_{RCA}$ is determined by current timing information through (xxx).
\begin{eqnarray}\label{CSA_RCA}
%  \mu_{c}\cdot n_{RCA}=\mu_{c}n_0+\mu_{mux}
  \mu_{c}\cdot n_{RCA}=d_i
\end{eqnarray}

Combining~(\ref{CSA_WL}) and~(\ref{CSA_RCA}), we derive the representation of the word-length of CSA in terms of a given timing constraint, as presented in~(\ref{CSA_Timing}).
% we form the relationship between the word-length of CSA and RCA under a given timing constraint, as presented in~(\ref{CSA_Timing}).
\begin{eqnarray}\label{CSA_Timing}
  n_{CSA}=s\cdot n_{RCA}-\frac{\mu_{mux}}{\mu_{c}}\cdot\frac{(s+2)(s-1)}{2}
\end{eqnarray}

@@

\subsection{Model Verification}
As seen in (\ref{CSA_Timing}), the value of $\mu_{mux}/\mu_c$ should be determined before applying the timing model. This ratio can be obtained through timing analysis instead of experiments. Specifically, varying the word-length of a single stage $n_i$ and recording the corresponding delay value $d_i$ from the timing analysis tool. According to~(\ref{CSA_SingleStageDelay}), the ratio can be obtained by fitting those values using linear polynomials.

Using this information, we verify our timing model with experimental results, which are obtained from post place and route simulations on Xilinx Virtex-6 FPGAs. Fig.~\ref{CSA Model Verification} illustrates both the modeled value and the experimental results of the maximum word-length of the 2-stage CSA for various operating frequencies. In the following experiments within this section, the input data are randomly sampled from a 16-bit data set, which follows uniform distribution. It can be seen that the the modeled value match well with the experimental results.
\begin{figure}[htbp]
  \centering
  \includegraphics[width=3in]{./Figures/Model.eps}
  \caption{Comparison of the maximum word-length for given operating frequencies between the modeled value and the experimental results from FPGA simulations.}
  \label{CSA Model Verification}
\end{figure}

@@

\subsection{Area Overhead}
We then compare different adder structures in terms of the maximum word-length for a given operating frequency and the corresponding area consumption. Fig.~\ref{CSA3stage Timing} illustrates the maximum word-lengths for RCA and CSA with 2 stages and 3 stages under a range of operating frequencies. We only investigate 3 stages as the maximum stages number predicted in (\ref{CSA_StageNumber}) is 3. However, similar analysis can be performed on CSA with any stage numbers and the overall conclusion will still hold. It can be seen from Fig.~\ref{CSA3stage Timing} that in comparison to RCA, CSA achieves greater word-length when frequency is initially increased. This corresponds to smaller truncation errors. RCA only outperforms than CSA when very high frequency is applied. In addition, the word-length of 3-stage CSA is always greater than 2-stage CSA across the entire frequency domain, as expected.
%\begin{figure}[htbp]
%  \centering
%  \includegraphics[width=3.3in]{./Figures/CSA3stage_Timing.eps}
%  \caption{Maximum word-length for RCA and CSA across a variety of frequencies.}
%  \label{CSA3stage Timing}
%\end{figure}

\begin{figure}[tbp]
    \begin{minipage}[b]{.48\textwidth}
        \centering
        \subfigure[Maximum word-length.]{
            \includegraphics[width=3.3in]{./Figures/CSA3stage_Timing.eps}
            \label{CSA3stage Timing}
        }
        \subfigure[Hardware resource usage.]{
            \includegraphics[width=3.3in]{./Figures/CSA3stage_Area.eps}
            \label{CSA3stage Area}
        }
    \end{minipage}
    \centering
    \caption{Comparison between RCA and CSA across a variety of frequency values in terms of the maximum word-length of input signal and the area consumption.}
\end{figure}

However, the accuracy benefits brought by CSA comes with the cost of large area overhead. Fig.~\ref{CSA3stage Area} depicts the number of Look-Up Tables (LUTs) in the FPGA used for all three structures. It can be seen that for all frequency values, the 3-stage CSA consumes $2.4\times\sim3.7\times$ area than RCA, while the number of the 2-stage CSA is $1.7\times\sim3.1\times$. This finding poses a question of whether CSA with higher stage numbers still offer better accuracy than RCA under a limited area budget.
%\begin{figure}[htbp]
%  \centering
%  \includegraphics[width=3.3in]{./Figures/CSA3stage_Area.eps}
%  \caption{Area overhead for RCA and CSA.}
%  \label{CSA3stage Area}
%\end{figure}

\subsection{Exploring Trade-offs Between Accuracy, Performance and Area}
To answer this question, the silicon area is incorporated into the trade-offs between accuracy and performance. If the available hardware resources are limited, the full word-length of both CSA and RCA might not be implemented. The precision loss might results in large errors even at low frequencies. This is demonstrated by experiments where area constraints are applied besides the timing requirement. In the conventional design scenario, both of the constraints are met by reducing the word-length of the input signal. The errors at the outputs are recorded with the reference to the original word-length, which is 16 bits in the following experiments. In addition, we propose another designing method where RCA is implemented with the maximum possible word-length under the given area budget, while the timing constraints are met by overclocking. For a certain pair of $\{Area, Frequency\}$ constraint, the optimum design method is selected based on the following criteria:
%For instance, Fig.xxx to Fig.xxx illustrate error expectations of all the described design scenarios when setting the LUTs number to 45, 35, 25 and 15 respectively. The optimal design metric with the minimum error expectation is labelled for each area constraint. For a given frequency and area constraint, the optimum design is chosen based on the following metrics:
\begin{itemize}
  \item Design with the minimum error expectation is the optimum design;
  \item If multiple designs achieve the same error expectation, then the design with minimum area is the optimum design;
  \item If the error expectation and area are identical for multiple designs, they are all treated as the optimum design.
\end{itemize}

The optimum design method for a variety of frequency and area constraints is demonstrated in Fig.~\ref{Tradeoff}, from which several observations can be made. If the available area is large enough, CSA serves as the optimum design option because it inherently operates faster than RCA, and it can be implemented with full precision with unlimited hardware resources. The 2-stage CSA is better than the 3-stage CSA when frequency is initially increased, as it consumes less area although both of them achieve the same error expectation. For a tighter area budget, only part of the CSA can be implemented, whilst RCA still keeps full precision. In this case, area instead of timing becomes the dominate factor. This leads to precision loss for CSA initially. Therefore we see RCA with overclocking is the optimum design method across almost the whole frequency domain. For an even stringent area constraint, the word-length of RCA is also limited. This results in truncation error initially for all design scenarios. Meanwhile, CSA with high stage numbers might not even be implemented at all.
\begin{figure}[htbp]
  \centering
  \includegraphics[width=3.5in]{./Figures/Tradeoff.eps}
  \caption{Demonstration of the optimum design methodology for a variety frequency and area constraints.}
  \label{Tradeoff}
\end{figure}
